Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-04-07 Cost_manuscript.Rmd
methods + results draft sans modeling

```{r Setups, include = FALSE}
## libraries
library(tidyverse)
library(ggfortify)
library(knitr)
library(pander) 
library(nloptr)
library(pwr)
library(lme4)
#library(lmerTest)
library(gridExtra)
library(reshape2)
library(corrplot)
library(survival)
library(ggfortify)

@@

Effort is costly, but sometimes it can add value. In thinking about this, studies have focused on TAFC. We have learned much about this (show data), but two issues: 1) many studies compare costs by comparing discounting functions, which is helpful, but 2) an often disregarded aspect of this is OC. Even in tafc, a choice for a sooner or easier item can be a reflection of wanting to skip the offer altogether. Foraging provides a natural way to probe this. Foraging is often divided into two: patch and prey, and they are often experienced in daily life. Patch leaving can be deciding when to leave a job (diminishing gains). Prey can be X, or nerflix (show gif).

@@

- I guess you can mention that the wait group acts as a baseline of foraging behavior.
- Re-read Mikey's effort value paper to see if there is anything there that can allow us to transition into foraging. Like, are there elements that make cog effort valuable that are well encapsuled by foraging paradigms?

@@

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and subject ID as a random effect (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in the following order: 1) intercept only; 2) condition only; 3) reward only; 4) condition and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

  Next we probed the confidence and consistency with which participants made choices. We computed the proportion of quits that were performed during the choice window versus during the handling time. We interpreted as confidence if a participant made at least 80% of their choices during the offer window. We then computed each participant’s total proportion of acceptances on the first two and last two blocks, and compared them using a paired permutation analysis (5000 iterations).
  **adapt once we decide on the model candidates**To estimate the subjective opportunity cost of each condition, we used a logistic function to model each participant’s probability of accepting a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. As before, OC was computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function were estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group were compared using a repeated measures ANOVA with cost as a factor. We performed pairwise paired permutations as post-hoc tests.

@@

``` {r BTW: 16.2.2. rmANOVA, echo=FALSE,fig.align="center",fig.width=5,fig.height=4, results = 'asis'}
# temp <- dataBtw %>%
#         filter(Choice < 2) %>%
#         group_by(SubjID, Cost, Handling, Offer) %>%
#         summarise(pAccept = mean(Choice))
# 
# ## REPEATED MEASURES ANOVA
# # I think I like this the most
# prop.aov <- with(temp, aov(pAccept ~ factor(Cost) * Handling * Offer + 
#                                  Error(SubjID / (Handling * Offer))))
# 
# ## NOTES ON RESULTS
# # According to the logistic fit, there are main effects for all predictors, but no interactions. It also says that the overall odds for the wait group to accept versus the effort is around 16%. Similarly, increasing the handling time results in a decrease in the multiplicative odds of ~25%. Increments in reward increase the odds of acceptance by ~40%.
```

@@

# and plot
(a1 <- dataBtw %>%
        filter(Choice < 2) %>%
        group_by(SubjID, Cost, Handling, Offer, optimal) %>%
        summarise(pAccept = mean(Choice)) %>%
        group_by(Cost, Handling, Offer, optimal) %>%
        summarise(meanComplete = mean(pAccept),
                  SE = sd(pAccept) / sqrt(length(pAccept))) %>%
   ggplot(aes(interaction(Offer, Handling), meanComplete, color = Cost)) + 
        geom_point(size = 3) + 
        geom_errorbar(aes(ymin = meanComplete - SE, ymax = meanComplete + SE), width = 0.2, size = 1) +
        geom_line(aes(group = interaction(Handling, Cost)), size = 1) +
        geom_point(aes(y = round(optimal)), shape = 21, fill = "grey75", color = "grey30", size = 3, show.legend = F) +
        labs(x = "Offer.Handling", y = "Proportion Accepted") +
        scale_color_manual(values = colsBtw) +
        theme(legend.key = element_blank(),
              panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              panel.background = element_blank(), 
              axis.line = element_line(colour = "black"),
              text = element_text(size = 16)))

@@

``` {r BTW: 16.2.4. Model comparison, echo=FALSE,fig.align="center",fig.width=6,fig.height=4}
# relevel, since the wait-reference model is used from the previous section
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Wait"))

@@

## Model selection 
# AIC/BIC
ME_evals <- list()
ME_evals$abicLogis_compare <- data.frame(Model = names(mixLogis_compare),
                                AIC = sapply(mixLogis_compare, AIC),
                                BIC = sapply(mixLogis_compare, BIC))

@@

``` {r BTW: 16.2.4. Plot Deviance comparison, echo=FALSE,fig.align="center",fig.width=6,fig.height=4}
# plot AIC and BIC
(mixLogis_compare$plot <- ME_evals$abicLogis_compare %>% 
  gather(Comparison, Value, AIC:BIC) %>%
  mutate(Model = factor(Model, levels = names(mixLogis_compare))) %>%
  ggplot(aes(Model, Value, fill = Comparison)) +
    geom_point(color = "black", pch = 21, size = 3) +
    theme_classic())

@@

survPlot <- autoplot(survData) +
  geom_vline(xintercept = 1, lty = 2) + # end of the grace period for physical trials
  scale_color_manual(values = colsBtw) +
  scale_fill_manual(values = colsBtw) +
  scale_x_continuous(breaks = seq(14)) +
  ylim(0, 1) +
  labs(x = "Time in Trial", y = "Proportion Accepted", fill = "Cost", color = "Cost") +
  theme(legend.position = c(0.35, 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 16))

# supplemental
# RT ecdfs for the whole group across blocks, with the option to subdivide by cost group
RTplot <- dataBtw %>%
  filter(Choice == 0,
         RT < 1) %>%
  ggplot(aes(RT, group = Block, color = Block)) +
    stat_ecdf() +
    labs(y = "Proportion of Quits Before Time X", x = "Quitting Time") +
    facet_wrap(vars(Cost)) +
    theme_classic()

grid.arrange(survPlot, RTplot, ncol = 2)

@@

## PRE/POST PAIRED PERMUTATIONS
temp <- dataBtw %>%
  filter(Choice < 2) %>%
  group_by(SubjID, Cost) %>%
  summarise(propAccept_1 = mean(Choice[Half == "Half_1"]),
            propAccept_2 = mean(Choice[Half == "Half_2"])) %>%
  plyr::dlply("Cost", identity)

# paired perms and effect sizes
prepost <- list()
prepost$CohenD <- sapply(temp, function(data) {DescTools::CohenD(data$propAccept_1, data$propAccept_2)})
prepost$Perms <- sapply(temp, function(data) {permute(data$propAccept_1, data$propAccept_2, paired = T, simple = T)})
```

@@

# and plot
prepost$scatterplot <- dataBtw %>%
                        filter(Choice < 2) %>%
                        group_by(SubjID, Cost) %>%
                        summarise(propAccept_1 = mean(Choice[Half == "Half_1"]),
                                  propAccept_2 = mean(Choice[Half == "Half_2"])) %>%
                        ggplot(aes(propAccept_1, propAccept_2, fill = Cost)) +
                          geom_point(size = 3, pch = 21, color = "black", show.legend = F) +
                          scale_fill_manual(values = colsBtw) +
                          scale_color_manual(values = colsBtw) +
                          xlim(0, 1) +
                          ylim(0, 1) +
                          labs(x = "Proportion Accepted - 1st Half", y = "Proportion Accepted - 2nd Half") +
                          geom_abline(slope = 1, intercept = 0, lty = 2) +
                          geom_hline(data = df_mean, aes(yintercept = propAccept_2, color = Cost), linetype="dashed", show.legend = F) +
                          geom_vline(data = df_mean, aes(xintercept = propAccept_1, color = Cost), linetype="dashed", show.legend = F) +
                          theme(panel.grid.major = element_blank(),
                                panel.grid.minor = element_blank(),
                                panel.background = element_blank(),
                                axis.line = element_line(colour = "black"),
                                text = element_text(size = 16))

# acceptances per cost across blocks
# note that because the same handling/travel time combos repeat in the same order pre/post, 1/4-2/4-3/6 are direct comparisons
prepost$propBlocks <- dataBtw %>%
                         filter(Choice < 2) %>%
                         group_by(SubjID, Cost, Block) %>%
                         summarise(propAccept_subj = mean(Choice)) %>%
                         group_by(Cost, Block) %>%
                         summarize(propAccept = mean(propAccept_subj),
                                   seAccept = sd(propAccept_subj)/sqrt(nSubjs_btw)) %>%
                         ggplot(aes(Block, propAccept, color = Cost)) +
                           geom_vline(xintercept = 3.5, linetype = "dashed") +
                           geom_point(size = 2) +
                           geom_errorbar(aes(ymin = propAccept - seAccept, ymax = propAccept + seAccept, color = Cost), width = 0.1) +
                           geom_line() +
                           ylim(0, 1) +
                           labs(y = "Proportion Acccepted") +
                           scale_x_continuous(breaks = seq(6)) +
                           scale_color_manual(values = colsBtw) +
                           scale_fill_manual(values = colsBtw) +
                           theme(legend.key = element_blank(),
                                 legend.position = c(0.8, 0.25),
                                 panel.grid.major = element_blank(),
                                 panel.grid.minor = element_blank(),
                                 panel.background = element_blank(),
                                 axis.line = element_line(colour = "black"),
                                 text = element_text(size = 16))

grid.arrange(prepost$scatterplot, prepost$propBlocks, ncol = 2)
```

@@

  Before computationally modeling behavior, we asked whether foragers had a stable representation of their environment. Since participants had complete a priori knowledge of the statistics of each environment, and were unaware of any other cost groups that could modulate the subjective cost of their assigned demand, we expected them to 1) show confidence in their choices by not quitting mid-handling time; and 2) adopt the same strategy while re-experiencing each environment. We investigated the first prediction by plotting a survival curve of trial quits across all participants (Figure X, left), with black crosses signaling either completed trials (at 2 s, 10 s, and 14 s) or forced travels. This qualitative analysis shows that most decisions to quit happened within the first second into the handling time, and participants rarely quit during the handling time. Physical and easy groups show slightly lagged responses because the experiment allowed a one-second grace period for participants to begin gripping (marked by the dashed line). This resulted in some individuals choosing to wait for that second to indicate an offer rejection (the censor at 4s in the physical group was a completion during a 2 s handling trial that was incorrectly recorded as 4 s by the experimental code). Plotting the quitting time distributions across blocks for the first handling second (ECDFs on Figure X, right) shows that participants in the wait and cognitive groups responded faster as the experiment progressed (note that participants in physical conditions mostly relied on letting the grace period expire). These response time observation  s suggest clarity in participant preferences.

  We next examined the second prediction (i.e. strategy stability) by comparing acceptances rates throughout the experiment. The scatter plot on Figure X shows the overall proportion accepted by each participant before and after the mid-point break (colored dashed lines indicate the mean for that group). Paired permutations on means of proportions showed no differences for wait and easy groups (*p* > 0.05), but significantly lower acceptance rates for physical (*p* = `r report_p(prepost$Perms[1])`, Cohen's D = `r round(prepost$CohenD[1], digits = 2)`) and cognitive (*p* = `r report_p(prepost$Perms[2])`, Cohen's D = `r round(prepost$CohenD[2], digits = 2)`) groups. The decline in acceptances appeared to be gradual for the cognitive group, but not for the physical group (Figure X, half point indicated by the dashed line). An explanation for these results is that these conditions became costlier due to fatigue, thus providing evidence that our cognitive and physical manipulations were effortful.

  It is possible that this reduction in acceptances affected the relative differences among conditions. We addressed this possibility by examining pre-post acceptance rates for each handling time and reward combination. Figure X shows the proportion of acceptances for each offer combination, with each half depicted on each column. This collage suggests that changes in choice behavior for effortful groups were mostly driven by the longest handling time, perhaps because of the fatigue suggested above. Regardless, the relative differences in acceptances among cost groups remain mostly untouched. We tested this observation by repeating the winning mixed-effects model from the previous section (i.e. all main effects) on each experimental half separately. We replicated the matrix of coefficients from the previous section, with each half mirrored along the diagonal (i.e. the top triangle shows the coefficients for the second half of the experiment). Together, these results suggest that decision makers had a strong representation of the environment by showing that 1) participants were confident in their choices and strategies; 2) cognitive and physical costs produced the expected fatigue from exerting effort; and 3) despite the resulting reduction of acceptances, the relative cost of each condition was consistent throughout the experiment.

@@

  This experiment probed whether the value-modulatory effects of each type of demand depended on the ability to obtain rewards through other behavioral means. If the cost or value of a demand are intrinsic, the relative preferences displayed above should be robust to alternatives. The experimental setup was mostly preserved, with 3 key differences: 1) all participants experienced cognitive effort, physical effort, and delay; 2) each block contained offers with one type of effort and passive waiting (allowing for comparisons in acceptances through the common wait condition). and 3) the handling and travel times were fixed at 10 s and 6 s, respectively. We therefore compared choices from the within-subjects experiment to the matching time condition from the between-subjects experiment. We hypothesized that reward decision patterns would match those of the previous experiment, and that interleaving effort and delay would produce persistent preferences for passive delay trials.

@@

  Analyses from the previous experiment were repurposed to examine cost differences within individuals. We begun by testing the prediction that rewards would positively influence choices. A Wilcoxon signed rank test showed that individual logistic coefficients for reward amount were significantly greater than 0 (mean $\beta$ = `r round(mean(baseCoeffs$Offer), digits = 2)`, SD = `r round(sd(baseCoeffs$Offer), digits = 2)`; V = `r wilcox.test(baseCoeffs$Offer)$statistic`, *p* `r report_p(wilcox.test(baseCoeffs$Offer)$p.value)`), once again confirming that participants were properly influenced by reward amounts. Adding a coefficient that tracked prior consecutive quits indicated that only `r sum(baseCoeffs$AR_effect)` individuals were influenced by recent quit history, thus discarding systematic sequential effects. In terms of earnings, a linear model predicting earnings once again showed that participants who accepted too few or too many trials earned the least, in line with the experimental design (Figure X; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits= 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, R-squared = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). When assessing the optimality of acceptances for 4 and 8 cents (99% of 20 cent offers were accepted), chi-squared tests against the optimal of 50% showed that participants under-accepted physical effort trials (proportion accepted = `r round(summary_costvsreward["Physical", "probability"], digits = 2)`, $\chi^2$ = `r round(summary_costvsreward["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"])`; all other *p* > 0.05). Despite this deviation, participants earned a similar monetary amount per cost on average (mean Cognitive = `r filter(earnings_within, Cost == "Cognitive")$mEarn`, SD = `r filter(earnings_within, Cost == "Cognitive")$sdEarn`; mean Wait-cognitive = `r filter(earnings_within, Cost == "Wait_C")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait_C")$sdEarn`; mean Physical = `r filter(earnings_within, Cost == "Physical")$mEarn`, SD = `r filter(earnings_within, Cost == "Physical")$sdEarn`; mean Wait-physical = `r filter(earnings_within, Cost == "Wait_P")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait_P")$sdEarn`). These results convey that participants in both experiments were similarly influenced by the characteristics of the environment, although participants in the present experiment gravitated more towards optimality (**big caveat: the same percentage can be produced by many non-optimal combinations, including accepting half of each reward. So neither approach really gets at this too well.**).

@@

# method 2
mixLogis_main <- list()
mixData <- dataWth %>%
  group_by(SubjID, Cost, Offer) %>%
  summarize(totalChoices = length(Choice),
            totalAccepted = sum(Choice),
            totalQuits = sum(Choice == 0),
            propAccepted = mean(Choice))

@@

>>>>

```{r Setups, include = FALSE}
## libraries
library(tidyverse)
library(ggfortify)
library(knitr)
library(pander) 
library(nloptr)
library(pwr)
library(lme4)
#library(lmerTest)
library(gridExtra)
library(reshape2)
library(corrplot)
library(survival)
library(ggfortify)
library(patchwork)

@@

Effort is costly, but sometimes it can add value (For the talk, use the T example with Angelina, but think about how well it fits in the "prefer effort"). In thinking about this, studies have focused on TAFC. We have learned much about this (show data), but two issues: 1) many studies compare costs by comparing discounting functions, which is helpful, but 2) an often disregarded aspect of this is OC. (other frame: What we've learned is how increasing a given demand reduces the value of its associated reward, but less explored is the question about how people choose among them, specifically how people allocate their time to engage with demands). Even in tafc, a choice for a sooner or easier item can be a reflection of wanting to skip the offer altogether. Foraging provides a natural way to probe this. Foraging is often divided into two: patch and prey, and they are often experienced in daily life. Patch leaving can be deciding when to leave a job (diminishing gains). Prey can be X, or nerflix (show gif).

@@

What we've learned is how increasing a given demand reduces the value of its associated reward, but less explored is the question about how people choose among them, specifically how people allocate their time to engage with demands.

- I guess you can mention that the wait group acts as a baseline of foraging behavior.
- Re-read Mikey's effort value paper to see if there is anything there that can allow us to transition into foraging. Like, are there elements that make cog effort valuable that are well encapsuled by foraging paradigms?
- Pending questions: explain logistic coefficients in exponentiated form? (i.e. X% more/less likely than...)

@@

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and subject ID as a random effect (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in the following order: 1) intercept only; 2) cost only; 3) reward only; 4) cost and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

  Next we probed the confidence and consistency with which participants made choices. First, we qualitatively assessed the number of quits performed during during the handling time through a survival curve, and plotted the quit response time distribution as a function of block to examine choice confidence over time. Next, we computed the proportion of quits that were performed during the choice window versus during the handling time. We interpreted as confidence if a participant made at least 80% of their choices during the offer window. We then computed each participant’s total proportion of acceptances on the first two and last two blocks, and compared them using a paired permutation analysis (5000 iterations). We avoided the third and fourth block because their effort designation was counterbalanced across participants, which would have biased our estimates.

  **adapt once we decide on the model candidates**To estimate the subjective opportunity cost of each condition, we used a logistic function to model each participant’s probability of accepting a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. As before, OC was computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function were estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group were compared using a repeated measures ANOVA with cost as a factor. We performed pairwise paired permutations as post-hoc tests.

@@

``` {r BTW: 16.2.2. rmANOVA, echo=FALSE,fig.align="center",fig.width=5,fig.height=4, results = 'asis'}
## This section was removed because it was deemed redundant. The interaction is addressed once the a priori mixed effects logistic model is evaluated against multiple forms of interaction effects (16.2.4.).
# temp <- dataBtw %>%
#         filter(Choice < 2) %>%
#         group_by(SubjID, Cost, Handling, Offer) %>%
#         summarise(pAccept = mean(Choice))
# 
# ## REPEATED MEASURES ANOVA
# # I think I like this the most
# prop.aov <- with(temp, aov(pAccept ~ factor(Cost) * Handling * Offer + 
#                                  Error(SubjID / (Handling * Offer))))
# 
# ## NOTES ON RESULTS
# # According to the logistic fit, there are main effects for all predictors, but no interactions. It also says that the overall odds for the wait group to accept versus the effort is around 16%. Similarly, increasing the handling time results in a decrease in the multiplicative odds of ~25%. Increments in reward increase the odds of acceptance by ~40%.
```

@@

# and plot
(hrPlot_btw <- dataBtw %>%
        filter(Choice < 2) %>%
        group_by(SubjID, Cost, Handling, Offer, optimal) %>%
        summarise(pAccept = mean(Choice)) %>%
        group_by(Cost, Handling, Offer, optimal) %>%
        summarise(meanComplete = mean(pAccept),
                  SE = sd(pAccept) / sqrt(length(pAccept))) %>%
   ggplot(aes(interaction(Offer, Handling), meanComplete, color = Cost)) + 
        geom_point(size = 3) + 
        geom_errorbar(aes(ymin = meanComplete - SE, ymax = meanComplete + SE), width = 0.2, size = 1) +
        geom_line(aes(group = interaction(Handling, Cost)), size = 1) +
        geom_point(aes(y = round(optimal)), shape = 21, fill = "grey75", color = "grey30", size = 3, show.legend = F) +
        labs(x = "Offer.Handling", y = "Proportion Accepted") +
        scale_color_manual(values = colsBtw) +
        theme(legend.key = element_blank(),
              panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              panel.background = element_blank(), 
              axis.line = element_line(colour = "black"),
              text = element_text(size = 16)))

@@

``` {r BTW: 16.2.4. Model comparison, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# relevel, since the wait-reference model is used from the previous section
mixData <- within(mixData, Cost <- relevel(Cost, ref = "Wait"))

@@

## Model selection 
# AIC/BIC
ME_evals <- list()
ME_evals$abicLogis_compare <- data.frame(Model = names(mixLogis_compare),
                                AIC = sapply(mixLogis_compare, AIC),
                                BIC = sapply(mixLogis_compare, BIC)) %>%
                                  mutate_at(vars(AIC:BIC), round)

@@

``` {r BTW: 16.2.4. Plot Deviance comparison, echo = FALSE, fig.align = "center", fig.width = 6,fig.height = 4}
# plot AIC and BIC
(mixLogis_compare$plot <- ME_evals$abicLogis_compare %>% 
  gather(Comparison, Value, AIC:BIC) %>%
  mutate(Model = factor(Model, levels = names(mixLogis_compare))) %>%
  ggplot(aes(Model, Value, fill = Comparison)) +
    geom_point(color = "black", pch = 21, size = 3) +
    theme_classic())

@@

survPlot_btw <- autoplot(survData) +
  geom_vline(xintercept = 1, lty = 2) + # end of the grace period for physical trials
  scale_color_manual(values = colsBtw) +
  scale_fill_manual(values = colsBtw) +
  scale_x_continuous(breaks = seq(14)) +
  ylim(0, 1) +
  labs(x = "Time in Trial", y = "Proportion Accepted", fill = "Cost", color = "Cost") +
  theme(legend.position = c(0.35, 0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 16))

# supplemental
# RT ecdfs for the whole group across blocks, with the option to subdivide by cost group
RTplot_btw <- dataBtw %>%
  filter(Choice == 0,
         RT < 1) %>%
  ggplot(aes(RT, group = Block, color = Block)) +
    stat_ecdf() +
    labs(y = "Proportion of Quits Before Time X", x = "Quitting Time") +
    facet_wrap(vars(Cost)) +
    theme_classic()

survPlot_btw | RTplot_btw

@@

## PRE/POST PAIRED PERMUTATIONS
temp <- dataBtw %>%
  group_by(SubjID, Cost) %>%
  summarise(propAccept_1 = mean(Choice[Half == "Half_1"]),
            propAccept_2 = mean(Choice[Half == "Half_2"])) %>%
  plyr::dlply("Cost", identity)

# paired perms and effect sizes
prepost_btw <- list()
prepost_btw$CohenD <- sapply(temp, function(data) {DescTools::CohenD(data$propAccept_1, data$propAccept_2)})
prepost_btw$Perms <- sapply(temp, function(data) {permute(data$propAccept_1, data$propAccept_2, paired = T, simple = T)})
```

@@

# and plot
prepost_btw$scatterplot <- dataBtw %>%
                        filter(Choice < 2) %>%
                        group_by(SubjID, Cost) %>%
                        summarise(propAccept_1 = mean(Choice[Half == "Half_1"]),
                                  propAccept_2 = mean(Choice[Half == "Half_2"])) %>%
                        ggplot(aes(propAccept_1, propAccept_2, fill = Cost)) +
                          geom_point(size = 3, pch = 21, color = "black", show.legend = F) +
                          scale_fill_manual(values = colsBtw) +
                          scale_color_manual(values = colsBtw) +
                          xlim(0, 1) +
                          ylim(0, 1) +
                          labs(x = "Proportion Accepted - 1st Half", y = "Proportion Accepted - 2nd Half") +
                          geom_abline(slope = 1, intercept = 0, lty = 2) +
                          geom_hline(data = df_mean, aes(yintercept = propAccept_2, color = Cost), linetype="dashed", show.legend = F) +
                          geom_vline(data = df_mean, aes(xintercept = propAccept_1, color = Cost), linetype="dashed", show.legend = F) +
                          theme(panel.grid.major = element_blank(),
                                panel.grid.minor = element_blank(),
                                panel.background = element_blank(),
                                axis.line = element_line(colour = "black"),
                                text = element_text(size = 16))

# acceptances per cost across blocks
# note that because the same handling/travel time combos repeat in the same order pre/post, 1/4-2/4-3/6 are direct comparisons
prepost_btw$propBlocks <- dataBtw %>%
                         filter(Choice < 2) %>%
                         group_by(SubjID, Cost, Block) %>%
                         summarise(propAccept_subj = mean(Choice)) %>%
                         group_by(Cost, Block) %>%
                         summarize(propAccept = mean(propAccept_subj),
                                   seAccept = sd(propAccept_subj)/sqrt(nSubjs_btw)) %>%
                         ggplot(aes(Block, propAccept, color = Cost)) +
                           geom_vline(xintercept = 3.5, linetype = "dashed") +
                           geom_point(size = 2) +
                           geom_errorbar(aes(ymin = propAccept - seAccept, ymax = propAccept + seAccept, color = Cost), width = 0.1) +
                           geom_line() +
                           ylim(0, 1) +
                           labs(y = "Proportion Acccepted") +
                           scale_x_continuous(breaks = seq(6)) +
                           scale_color_manual(values = colsBtw) +
                           scale_fill_manual(values = colsBtw) +
                           theme(legend.key = element_blank(),
                                 legend.position = c(0.8, 0.25),
                                 panel.grid.major = element_blank(),
                                 panel.grid.minor = element_blank(),
                                 panel.background = element_blank(),
                                 axis.line = element_line(colour = "black"),
                                 text = element_text(size = 16))

prepost_btw$scatterplot | prepost_btw$propBlocks
```

@@

  Before computationally modeling behavior, we asked whether foragers had a stable representation of their environment. Since participants had complete a priori knowledge of the statistics of each environment, and were unaware of any other cost groups that could modulate the subjective cost of their assigned demand, we expected them to 1) show confidence in their choices by not quitting mid-handling time; and 2) adopt the same strategy while re-experiencing each environment. We investigated the first prediction by plotting a survival curve of trial quits across all participants (Figure X, left), with black crosses signaling either completed trials (at 2 s, 10 s, and 14 s) or forced travels. This qualitative analysis shows that most decisions to quit happened within the first second into the handling time, and participants rarely quit during the handling time. Physical and easy groups show slightly lagged responses because the experiment allowed a one-second grace period for participants to begin gripping (marked by the dashed line). This resulted in some individuals choosing to wait for that second to indicate an offer rejection (the censor at 4s in the physical group was a completion during a 2 s handling trial that was incorrectly recorded as 4 s by the experimental code). Plotting the quitting time distributions across blocks for the first handling second (ECDFs on Figure X, right) shows that participants in the wait and cognitive groups responded faster as the experiment progressed (note that participants in physical conditions mostly relied on letting the grace period expire). These response time observations suggest clarity in participant preferences.

  We next examined the second prediction (i.e. strategy stability) by comparing acceptances rates throughout the experiment. The scatter plot on Figure X shows the overall proportion accepted by each participant before and after the mid-point break (colored dashed lines indicate the mean for that group). Paired permutations on means of proportions showed no differences for wait and easy groups (*p* > 0.05), but significantly lower acceptance rates for physical (*p* = `r report_p(prepost_btw$Perms[1])`, Cohen's D = `r round(prepost_btw$CohenD[1], digits = 2)`) and cognitive (*p* = `r report_p(prepost_btw$Perms[2])`, Cohen's D = `r round(prepost_btw$CohenD[2], digits = 2)`) groups. The decline in acceptances appeared to be gradual for the cognitive group, but not for the physical group (Figure X, half point indicated by the dashed line). An explanation for these results is that these conditions became costlier due to fatigue, thus providing evidence that our cognitive and physical manipulations were effortful.

  It is possible that this reduction in acceptances affected the relative differences among conditions. We addressed this possibility by examining pre-post acceptance rates for each handling time and reward combination. Figure X shows the proportion of acceptances for each offer combination, with each half depicted on each column. This collage suggests that changes in choice behavior for effortful groups were mostly driven by the longest handling time, in agreement with previous findings where increments in effort duration correlated with aversion (Treadway et al., 2009), and with the fatigue suggested above. Regardless, the relative differences in acceptances among cost groups remain mostly untouched. We tested this observation by repeating the winning mixed-effects model from the previous section (i.e. all main effects) on each experimental half separately. We replicated the matrix of coefficients from the previous section, with each half mirrored along the diagonal (i.e. the top triangle shows the coefficients for the second half of the experiment). Together, these results suggest that decision makers had a strong representation of the environment by showing that 1) participants were confident in their choices and strategies; 2) cognitive and physical costs produced the expected fatigue from exerting effort; and 3) despite the resulting reduction of acceptances, the relative cost of each condition was consistent throughout the experiment.

@@

  This experiment probed whether the value-modulatory effects of each type of demand depended on the ability to obtain rewards through other behavioral means. If the cost or value of a demand are intrinsic, the relative preferences displayed above should be robust to alternatives. The experimental setup was mostly preserved, with 3 key differences. First, all participants experienced cognitive effort, physical effort, and delay. Second, each block contained offers with one type of effort and passive waiting (allowing for comparisons in acceptances through the common wait condition). We refer to observations from each block's wait trials by their effort pair (i.e. "Wait_C" for cognitive, and "Wait_P" for physical). Finally, in this experiment the handling and travel times were fixed at 10 s and 6 s, respectively. We therefore compared choices from the within-subjects experiment to the matching time condition from the between-subjects experiment. We hypothesized that reward decision patterns would match those of the previous experiment, and that interleaving effort and delay would produce persistent preferences for passive delay trials.

@@

  Analyses from the previous experiment were repurposed to examine cost differences within individuals. We begun by testing the prediction that rewards would positively influence choices. A Wilcoxon signed rank test showed that individual logistic coefficients for reward amount were significantly greater than 0 (mean $\beta$ = `r round(mean(baseCoeffs$Offer), digits = 2)`, SD = `r round(sd(baseCoeffs$Offer), digits = 2)`; V = `r wilcox.test(baseCoeffs$Offer)$statistic`, *p* `r report_p(wilcox.test(baseCoeffs$Offer)$p.value)`), once again confirming that participants were properly influenced by reward amounts. Adding a coefficient that tracked prior consecutive quits indicated that only `r sum(baseCoeffs$AR_effect)` individuals were influenced by recent quit history, thus discarding systematic sequential effects. In terms of earnings, a linear model predicting earnings once again showed that participants who accepted too few or too many trials earned the least, in line with the experimental design (Figure X; general linear model with quadratic term, F = `r round(summary(lmEarn)$fstatistic[1], digits = 2)`, Beta = `r round(summary(lmEarn)$coefficients[3,1], digits= 2)`, SE = `r round(summary(lmEarn)$coefficients[3, 2], digits = 2)`, R-squared = `r round(summary(lmEarn)$r.squared, digits = 2)`, *p* `r report_p(summary(lmEarn)$coefficients[3, 4])`). When assessing the optimality of acceptances for 4 and 8 cents (99% of 20 cent offers were accepted), chi-squared tests against the optimal of 50% showed that participants under-accepted physical effort trials (proportion accepted = `r round(summary_costvsreward["Physical", "probability"], digits = 2)`, $\chi^2$ = `r round(summary_costvsreward["Physical", "chisquared"], digits = 2)`, *p* `r report_p(summary_costvsreward["Physical", "pval"])`; all other *p* > 0.05). Despite this deviation, participants earned a similar monetary amount per cost on average (mean Cognitive = `r filter(earnings_within, Cost == "Cognitive")$mEarn`, SD = `r filter(earnings_within, Cost == "Cognitive")$sdEarn`; mean Wait_C = `r filter(earnings_within, Cost == "Wait_C")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait_C")$sdEarn`; mean Physical = `r filter(earnings_within, Cost == "Physical")$mEarn`, SD = `r filter(earnings_within, Cost == "Physical")$sdEarn`; mean Wait_P = `r filter(earnings_within, Cost == "Wait_P")$mEarn`, SD = `r filter(earnings_within, Cost == "Wait_P")$sdEarn`). These results convey that participants in both experiments were similarly influenced by the characteristics of the environment, although participants in the present experiment gravitated more towards optimality (**big caveat: the same percentage can be produced by many non-optimal combinations, including accepting half of each reward. So neither approach really gets at this too well.**).

@@

# mixed effects logistics, rotating the reference cost condition
mixLogis_main <- list()
mixData <- dataWth %>%
  group_by(SubjID, Cost, Offer) %>%
  summarize(totalChoices = length(Choice),
            totalAccepted = sum(Choice),
            totalQuits = sum(Choice == 0),
            propAccepted = mean(Choice))

@@

# analysis of deviance
mixLogis_compare$aod <- anova(mixLogis_compare$Reward, mixLogis_compare$AllMain, mixLogis_compare$Interaction)
